---
title: "Milestone"
author: "ClarenceChen"
date: "7/21/2020"
output: html_document
---
## Introduction

This Milestone Report goes over the exploratory data analysis for the Capstone Project of the Data Science Specialization on Coursera

The Key partners for this project are Swiftkey and Coursera. The project explores the Natural Language Processing facet of Data Science

A large text corpus of documents to predict the next word on preceding input. 

After the initial extraction and cleaning of data, it is presented on a Shiny application

This report goes over the plan on data preparation and presentation from a large corpus of text

## Load R Libraries

Libraries used for this project.

```{r loading R libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(openxlsx)
library(ggplot2)
library(stringi)
library(NLP)
library(tm)
library(SnowballC)
library(RColorBrewer)
```

## Load The Data 
The data for this project is selected from HC Corpora

 * en_US.blogs.txt
 * en_US.news.txt
 * en_US.twitter.txt.

The data link is provided by Coursera. After downloading the data to a local machine, it is read into R, first with an open binary connection and then creating a variable for each set


```{r Loading the data, echo=TRUE, message=FALSE, warning=FALSE}
blogs <- file("en_US.blogs.txt", open="rb")
blogs <- readLines(blogs, encoding = "UTF-8", skipNul=TRUE)
news <- file("en_US.news.txt", open = "rb") 
news <- readLines(news, encoding = "UTF-8", skipNul=TRUE)
twitter <- file("en_US.twitter.txt", open = "rb")
twitter <- readLines(twitter, encoding = "UTF-8", skipNul=TRUE)
```

## Metadata

The loaded from each of the three files - blogs, news and twitter is now evaluated to check it's metadata. This mainly includes the size (in megabytes), number of lines in each variable (document), number of words, length of the longest line etc. 

### File Size
```{r Metadata, echo=TRUE}
file.info("en_US.blogs.txt")$size / 1024^2
file.info("en_US.news.txt")$size  / 1024^2
file.info("en_US.twitter.txt")$size / 1024^2
```

### Number of Lines
```{r Number of lines, echo=TRUE}
length(blogs)
length(news)
length(twitter)
```

### Word Count
```{r Word count, echo=TRUE}
sum(stri_count_words(blogs))
sum(stri_count_words(news)) 
sum(stri_count_words(twitter))
```

### The length of the longest line 
```{r Length of longest line, echo=TRUE}
max(nchar(blogs))
max(nchar(news))
max(nchar(twitter))
```


### Data Summary

* File size: All three files are over 200 MB
* Word Count: There are more than 30 million words in all the files
* As anticipated Twitter is a large file wil less words per line (140) and more lines
* CharaThe longest line among the blogs is 40,833 characters.
* News have longer paragraphs compared to Blogs and Tweets
